\section{Project Plan (\textit{Marcin})}
\label{sec:project_plan}

\subsection{Development Process}

To develop the language and its translator we were meeting at least once a 
week and were using software / project management tools to enhance the
development process. As a group we have met 13 times, and our team has also
met 7 times with the group's mentor, Prof. Alfred Aho. During the meetings 
we were brainstorming how Swift Fox should be designed and implemented.
Moreover, we were ensuring that we are on the right track to meet the
milestone submissions, and if necessary we were rescheduling the roles and 
assignments to meet the deadlines. The rule of thumb which was kept by the 
project manager was to have draft / alpha version of a document, or a code,
to be ready a week before the milestone date set by the instructor. Although this
last week before submission was expected to be used for polishing the final
work, corrections, and debugging, it was also used to finish the work which
was missed by other members of the group. 

To improve work assignment, keep work schedule on time, and improve
software engineering practice, a dedicated server with a web-based project
management and bug-tracking tool was setup on the Network Security
Research Lab (NSL) \cite{nsl}. Vasileios has dedicated a server and
installed Trac \cite{trac}. Trac is a project management tool with wiki
interface, ticket scheduling system, and source code repository. All group 
members received accounts on the server and have access to the Trac
management portal through the address
\url{http://nslvm2.cs.columbia.edu/trac/wiki}. After every meeting, job
tickets were assigned and all members could track the progress of the team 
work. Unfortunately, after a month the excitement from using Trac has
fallen down, and some members stopped using the tool.

\subsection{Roles and Responsibilities}

The group was started with three members: Marcin, Vasileios, and Linh.
After Yiwei joined us, we assigned the roles, which in practice were often 
reassigned to meet various milestones.

\begin{itemize}
	\item Project manager - Marcin (Vasileios was the second project
	manager)\\Marcin kept the project schedule, hold weekly meetings
	with the entire team, maintained the project log, and made sure
	the project deliverables get done on time
	\item Language and tools guru - Vasileios\\
	Vasileios defined the baseline process to track language changes
	and maintained the intellectual integrity of the language. Moreover
	he was teaching the team how to utilize the various tools used to
	build the compiler
	\item System architect - Marcin\\
	Based on Fennec Fox platform, Marcin defined the compiler
	architecture, modules, and interfaces
	\item System integrator - Yiwei\\
	For system integration, Yiwei designed the syntax tree traversal
	subsystem, which was used in the semantic check and code generation 
	steps
	\item Tester and validator - Linh (Vasileios was actually designing
	and coding the test suit)\\
	Linh was attempting to create the testing suit
\end{itemize}

\subsection{Implementation}

Swift Fox documentation was implemented using \LaTeX  and the code was
implemented in C. Both documentation and code were stored in the svn \cite{svn}
repository and managed using Trac \cite{trac} project management tool.

\begin{enumerate}
  \item Swift Fox Whitepaper
    \begin{itemize}
      \item Introduction and Problem Statement\\\textit{Marcin and Vasileios}
      \item Buzzwords: simple, declarative, domain specific, lightweight\\\textit{Linh}
      \item Buzzwords: network-savvy, platform independent, event-driven\\\textit{Marcin}
      \item Buzzwords: robust, secure, policy-oriented\\\textit{Vasileios}
      \item Draft Review\\\textit{Marcin, Vasileios, Yiwei}
    \end{itemize}

  \item Swift Fox Tutorial \& Manual
    \begin{itemize}
      \item Tutorial\\\textit{Marcin}
      \item Manual\\\textit{Vasileios}
      \item Draft Review\\\textit{Marcin and Vasileios}
    \end{itemize}

  \item Class Presentation
    \begin{itemize}
     \item Graphics and Text\\\textit{Marcin and Vasileios}  
     \item Implementing Presentation in \LaTeX\\\textit{Vasileios}
     \item Practice Presentation Meeting\\\textit{Marcin, Vasileios, Linh, Yiwei}
    \end{itemize}

  \item Swift Fox Compiler Code
    \begin{itemize}
      \item Lexer\\\textit{Marcin and Vasileios}
      \item Parser\\\textit{Marcin}
      \item Code Generation\\\textit{Marcin}
      \item Test Suit\\\textit{Vasileios}
      \item Semantic Check\\\textit{Vasileios}
      \item Tree Traversal\\{Yiwei}
    \end{itemize}

  \item Final Report
    \begin{itemize}
      \item Chapter 4\\\textit{Marcin}
      \item Chapter 5\\\textit{Vasileios}
      \item Chapter 6\\\textit{Marcin}
      \item Chapter 7\\\textit{Yiwei}
      \item Chapter 8\\\textit{Vasileios}
      \item Chapter 9\\\textit{Marcin, Vasileios, Linh, Yiwei}
      \item Draft Review\\\textit{Marcin, Vasileios, Linh}
    \end{itemize}
\end{enumerate}


\subsection{Timeline and Project Log}

The timeline of the project is listed below.

\begin{itemize}
\item January 28 - team formed by Marcin and Vasileios
\item February 9 - roles assigned for Marcin, Vasileios, and Linh (one member is missing)
\item February 22 - first meeting with all members present (Yiwei joins the team)
\item February 24 - Swift Fox white-paper was submitted 
\item March 12 - grammar, tokens, and types are finished
\item March 24 - Swift Fox tutorial and Swift Fox manual are submitted
\item April 12 - presentation of a simple swift fox example; blink demo works
\item April 25 - practice for the class presentation
\item April 26 - project presentations in class
\item May 6 - Swift Fox compiles more sophisticated programs than blink 
\item May 9 - report and code are submitted
\item May 10 - demo presentation
\end{itemize}

A more detailed, day-by-day timeline can be found at 
\url{http://nslvm2.cs.columbia.edu/trac/timeline}
The day-by-day timeline generated by Track shows every member contributions
to the project, including coding and writing text. In the Trac we were also
storing the meeting log. The meeting log showing all meetings with each
meeting's agenda can be found at
\url{http://nslvm2.cs.columbia.edu/trac/wiki/MeetingsLog}. Overall, our
group had 20 meetings, out of which 7 meetings were in the presence of the 
Prof. Alfred Aho. Marcin and Vasilis were present on all meetings, and Linh
and Yiwei both missed 6 meetings. Finally, the Trac system was generating
reports from all jobs done based on the assignment tickets assigned after each
meeting. Detailed tickets assignment log can be found at
\url{http://nslvm2.cs.columbia.edu/trac/report/6}.
